---
title: "Structural Equation Modelling class"
output: html_notebook
---

Close the back-door:

If we want to know the causal 

## 2nd day

correlation = if coefficients are standarized
covariance = if coefficients are not standarized

Standarize = demean and divided by SD.


```{r loading_libs}
# devtools::install_github("jslefsche/piecewiseSEM@devel")

library("lavaan")
library("piecewiseSEM")
library(tidyverse)


## load data
data(keeley)

keeley

skimr::skim(keeley)
```

First fit a linear model

```{r}
keeley_lm <- lm(cover ~ age, data = keeley)
summary(keeley_lm)
```

```{r}
# new model with lavaan
# lavaan requires the formula to be a character string
keeley_formula1 <- "cover ~ age"

#then pass to function 'sem'
keeley_sem1 <- sem(keeley_formula1, data = keeley)

summary(keeley_sem1)

## if you want to see the intercept
summary(sem(keeley_formula1, data = keeley, meanstructure = T))
```

t-rule to our model: 3 knows (2 variances, 1 covarance) - 3 estimated ($\gamma_{1,1}, \zeta_1, \delta_1$)= 0

The function does not return the standarized coeficients by default, one needs to use `coef`

```{r}
# get standardized coefficients of LM
coefs(keeley_lm)

# get standardized coefficeients of SEM using argument standardized = T in summary or the function 'standardizesoluion'
summary(keeley_sem1, standardized = T)
standardizedsolution(keeley_sem1)

# to get the R-squared, use the argument `rsq = T` in summary
summary(keeley_sem1, rsq = TRUE)

# compare to output from LM
summary(keeley_lm)$r.squared
```

## Next model

```{r}
formula2 <- "
  firesev ~ age
  cover ~ age + firesev"

sem2 <- sem(formula2, data = keeley)
summary(sem2, standardize = TRUE, rsq = TRUE)


```

Computing indirect effects:
```{r}
# in lavaan, can use identifiers for model coefficients int he fomrula
formula2.1 <- "
  firesev ~ af * age
  cover ~ ac * age + fc * firesev

  indirect := af * fc # := opeerator adds coefficients
  total := ac + af * fc"

summary(sem(formula2.1, keeley), standardize = TRUE)

```

### Exercise 1

```{r}
formula3 <- "
  rich ~ ar * abiotic + dr * distance + hr * hetero
  abiotic ~ da* distance
  hetero ~ dh* distance

  indirect1 := ar * da
  indirect2 := dh * hr
  total := dr + ar * da + dh * hr
"

sem3 <- sem(formula3, data = keeley)

summary(sem3, standardize = TRUE)
```


## Some visualization

```{r}
g <- standardizedsolution(sem3) %>% 
    filter(op == "~") %>% 
    select(-op) %>%
    select(rhs, lhs, est.std, se, z, pvalue) %>%
    igraph::graph_from_data_frame()

plot(g,
     edge.color = c("red", "black")[1 + (igraph::E(g)$est.std > 0)],
     edge.lty = c("solid", "dashed")[1 + (igraph::E(g)$pvalue > 0.05)],
     edge.label = round(igraph::E(g)$est.std, 2),
     edge.width = igraph::E(g)$est.std)

```

